\section{Introduction}

The \textbf{Three-Distance Theorem}---also called the \emph{three-gap theorem},
the \emph{Steinhaus theorem}, or the \emph{Sós--Surányi theorem}---is one of the
most elegant results in the metric theory of numbers.
It was conjectured by Hugo Steinhaus around 1957 and proved independently
by Vera Sós, János Surányi, and S.\,Świerczkowski in 1958.

\subsection*{The Basic Setup}

Fix a real number $\alpha$ and a positive integer $n$.
Place $n$ points on the unit circle $\mathbb{T} = \mathbb{R}/\mathbb{Z}$ at positions
\[
  \{{\alpha}\},\quad \{2\alpha\},\quad \ldots,\quad \{n\alpha\},
\]
where $\{x\} = x - \lfloor x \rfloor$ denotes the fractional part of $x$.
These $n$ points partition $\mathbb{T}$ into exactly $n$ arcs (gaps).

\subsection*{The Surprising Answer}

No matter which $\alpha$ and $n$ you choose, those $n$ arcs have
\textbf{at most three distinct lengths}.
Moreover, when exactly three lengths occur, the longest equals the sum of the other two.

This is remarkable: one might expect the number of distinct gap lengths to grow with $n$,
or to depend heavily on the arithmetic of $\alpha$.
Instead it is universally bounded by $3$.

\subsection*{Why It Matters}

The three-distance theorem sits at the junction of several fields:
\begin{itemize}
  \item \textbf{Uniform distribution.}
    The sequence $(\{n\alpha\})$ is equidistributed modulo $1$ for irrational $\alpha$
    (Weyl, 1916). The three-distance theorem gives quantitative control on \emph{how}
    the points spread out at finite stages.

  \item \textbf{Continued fractions.}
    The values of $n$ at which the gap structure transitions between two and three
    lengths are precisely the denominators of the convergents of $\alpha$,
    tying the theorem intimately to the Euclidean algorithm.

  \item \textbf{Quasicrystals and tilings.}
    One-dimensional quasicrystalline sequences (Beatty sequences, Sturmian words)
    exhibit exactly the three-distance structure; the theorem is the combinatorial
    backbone of their aperiodic order.

  \item \textbf{Music theory.}
    Equal temperament divides the octave into 12 steps, but composers have long
    used scales that divide it unequally. The three-distance theorem governs
    the gap structure of any such division.
\end{itemize}
