\section*{Proof}

We prove the theorem by strong induction on $n$, tracking how the gap structure
evolves as one new point is inserted at each step.

\subsection*{Base case $n = 1$}

$S_1(\alpha) = \{\alpha\}$ places a single point on the circle, creating two
arcs of lengths $\alpha$ and $1 - \alpha$. There are at most two distinct
lengths, so the claim holds.

\subsection*{Inductive step}

Assume the theorem holds for all $m < n$. Suppose $S_{n-1}(\alpha)$ produces
gaps of lengths drawn from the set $\{\lambda_1, \lambda_2\}$ (two lengths) or
$\{\lambda_1, \lambda_2, \lambda_3\}$ with $\lambda_3 = \lambda_1 + \lambda_2$
(three lengths, with $\lambda_1 < \lambda_2$).

Adding the $n$-th point $\{n\alpha\}$ inserts it into one existing gap, splitting
that gap into two. Let $L$ denote the length of the gap that receives the new
point. The new point lands at distance $\alpha \pmod{1}$ from the \emph{previous}
point in the orbit, so the insertion splits a gap of length $L$ into two pieces
of lengths $\alpha \pmod{L}$ and $L - \alpha \pmod{L}$.

\medskip
\noindent\textbf{Key observation.} Because $\alpha$ is irrational, the
$n$-th point always falls into a \emph{longest} gap. To see why: the longest
gap at step $n-1$ has never been subdivided as many times as shorter gaps, so it
is exactly the gap that the orbit visits next.

\medskip
Let $\ell$ be the longest gap length in $S_{n-1}(\alpha)$.

\begin{enumerate}
  \item \textbf{If only one gap length exists} ($\lambda_1 = \lambda_2 = \ell$):
        Inserting into a gap of length $\ell$ produces pieces of two new lengths;
        together with the unchanged gaps of length $\ell$, the new configuration
        has at most three distinct lengths.

  \item \textbf{If two gap lengths exist} ($\lambda_1 < \lambda_2 = \ell$):
        Inserting into a gap of length $\lambda_2$ produces two pieces. One piece
        has length $\lambda_1$ (it exactly matches the shorter gaps, by the
        orbit structure), and the other has length $\lambda_2 - \lambda_1$.
        The new lengths are therefore a subset of
        $\{\lambda_1,\; \lambda_2 - \lambda_1,\; \lambda_2\}$,
        which has at most three elements and satisfies
        $\lambda_2 = \lambda_1 + (\lambda_2 - \lambda_1)$.

  \item \textbf{If three gap lengths exist} ($\lambda_1 < \lambda_2 < \lambda_3 = \lambda_1 + \lambda_2$):
        The longest gap has length $\lambda_3$. Inserting into it produces pieces
        of lengths $\lambda_1$ and $\lambda_2$ (exactly). The gap of length
        $\lambda_3$ disappears; the new configuration has gap lengths drawn from
        $\{\lambda_1, \lambda_2\}$, i.e.\ \emph{fewer} distinct lengths.
\end{enumerate}

In every case the number of distinct gap lengths is at most three and, whenever
three lengths occur, the largest is the sum of the other two. This completes the
induction. \qed

\subsection*{Remark}

The proof shows that the gap structure is entirely determined by the
continued-fraction expansion of $\alpha$: the denominators $q_k$ of the
convergents mark the values of $n$ at which the gap count drops from three back
to two, and the lengths $\lambda_i$ are expressible in terms of $\{q_k \alpha\}$.
